\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-05-11 paper/paper.tex
More paper edits (Model Comparison) and updated references bibtek.

\subsection{Background}
In engineering applications, a product can often fail due to one out of a set of possible malfunctioning components.  For example, a computer system can fail if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of normal wearout.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \citet{nelson}.  A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all $s$ risks.  Parameter estimation is straightforward if the cause of failure is known for each observation.  With engineering systems data, however, the exact cause of failure is frequently unknown or masked from the researcher.  \\

Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.  When modeling computer system failures Reiser et al. assumed the cause of failure for each observation came from a known subset of failure modes, and estimation was performed using a Bayesian approach \cite{reiser}.  Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, Basu et. al used a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.  Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.  Treating the unknown failure modes as incomplete data, an expectation maximization algorithm with ML was used, in addition to Bayesian estimation.

\subsection{Motivation}
The goal of this paper is to model, and compare, the lifetime distribution of different hard drive brands using the GFLP lifetime model.  Product populations often contain a mixture of defective and non defective units.  The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.  Ignoring this hazard structure, which exists for series systems and computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.  The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.  Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are masked, which presents challenges for parameter identification: especially with sample sample sizes. One option is to combine all drive models and estiamte an overall GFLP model.  While this model can be precisely estimated, the data indicate a common model for all drive models fits well for only a few drive models.  At the other extreme, fitting individual models to each hard drive brand excludes drive brand models with one or two failures.  The GFLP model applied to these data produces very imprecise parameter estimates or can only be fit if highly informative priors are imposed.  Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.  With over 60 brand drive models in testing a hierarchical model is advantageous as it pools information from across brands to get more precise estimates of individual drive model parameters.  Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive models with lots of failures can be estimated precisely; and in cases where little data is available the hierarchical model borrows information from other models to obtain stable estimates.  This allows us to include all of the hard drive data and more importantly, make reliability inference for all the drive models in the data set.

@@

In addition to graphically examining model fits via posterior credible bands, we compared the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.  The predictive accuracy from a fitted Bayesian model can be estimated simply with posterior simulations of the model parameters \cite{vehtari}.  For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior parameters with the $i$th data point removed.  The final expected point-wise predictive density is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).  To obtain a more stable estimate of elpd, Vehtari et. al suggest using Pareto smoothed importance weights (PSIS) since the tail of the importance weights used to estimate elpd is often long.  We computed the elpd for all 4 models using the R package loo \cite{loopackage}.  When $n$ is large, the distribution of the elpd is approximately normal and models can be statistically compared.  We calculated the difference in expected predictive accuracy for all 4 models as well as the standard error of the difference.  As the model complexity increased, the predictive accuracy improved (Table 1).  Of all the models, Model 4 had the best predictive accuracy and was significantly better than models 1-3.

>>>>

\subsection{Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can fail if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \citet{nelson}.  A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all $s$ risks.  Parameter estimation is straightforward if the cause of failure is known for each observation.  With engineering systems data, however, the exact cause of failure is frequently unknown or masked from the researcher.  \\

Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.  When modeling computer system failures Reiser et al. assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach \cite{reiser}.  Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, Basu et. al performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.  Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.  Treating the unknown failure modes as incomplete data, an expectation maximization algorithm with ML was used, in addition to Bayesian estimation.

\subsection{Motivation}
The goal of this paper is to model, and compare, the lifetime distribution of different hard drive brands using the GFLP lifetime model.  Product populations often contain a mixture of defective and non defective units.  The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.  Ignoring this hazard structure, which exists for series systems, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{sujata}.  The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.  Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are masked, which presents challenges for parameter identification: especially with sample sample sizes. One option is to combine all drive models and estiamte an overall GFLP model.  While this model can be precisely estimated, the data indicate a common model for all drive models fits well for only a few drive models.  At the other extreme, fitting individual models to each hard drive brand excludes drive brand models with one or two failures.  The GFLP model applied to these data produces very imprecise parameter estimates or can only be fit if highly informative priors are imposed.  Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.  With over 60 brand drive models in testing a hierarchical model is advantageous as it pools information from across brands to get more precise estimates of individual drive model parameters.  Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive brand models to obtain stable estimates.  This allows us to include all of the hard drive data and more importantly, make reliability inference for all the drive models in the data set.

@@

In addition to graphically examining model fits via posterior credible bands, we compared the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.  As outlined in Vehtari et. al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters \citet{vehtari}.  For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with the $i$th data point removed.  The final expected point-wise predictive density is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).  To obtain a more stable estimate of elpd, Vehtari et. al suggest using Pareto smoothed importance weights (PSIS) since the tail of the importance weights used to estimate elpd is often long.  We computed the elpd for all 4 models using the R package loo \citet{loo}.  When $n$ is large, the distribution of the elpd is approximately normal and models can be statistically compared.  We calculated the difference in expected predictive accuracy for Model 4 vs 3, Model 3 vs 2, and Model 2 vs 1, as well as the standard error of the difference (Table 1).  As the model complexity increased, the expected predictive accuracy improved.  Of all the models, Model 4 had the best predictive accuracy and was significantly better than the other 3 models.

\begin{table}[H]
\centering
\begin{tabular}{rrrrrrr}
  \hline
 & elpd\ & Diff(elpd)\ & SE(Diff) \\ 
  \hline
Model 4 & -13309.5 & 40.7 & 11.3 \\ 
Model 3 & -13350.2 & 458.8 & 31.0  \\ 
Model 2 & -13809.0 & 3674.6 & 96.2 \\ 
Model 1 & -17483.6  \\ 
   \hline
\end{tabular}
\caption{Model Comparison Based on LOO}
\label{table:1}
\end{table}
